---
title: 'Homework 4: Data Import and Enhanced Graphics'
author: "NAME"
date: "DATE"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
```

# Importing data
Go to [Dr. D's data website](https://www.norcalbiostat.com/teaching/data/#) to find the following data files. Download them to your computer (ref: [HW 09](../09_import.html)) and import them in the first three questions below. 

1. Read in the `High School and Beyond` data set.

```{r}
hbs2 <- read.delim("../data/hsb2.txt", header = TRUE, sep = "\t")
```

2. Read in the `Crime Data` data set. 

```{r}
library(readxl)
crime <- read_excel("../data/Crime_Data.xlsx", sheet = 1, col_names = TRUE)
```

3. Read in the `countyComplete` data set.

```{r}
county <- read.csv("../data/countyComplete.csv", header = TRUE, sep = ",")
```

4. When reading in the Excel data set on crime, what do the arguments `sheet=1` and `col_names=TRUE` mean? 

Read from the first sheet and the first row contains column names.

5. Use the `str()` function to determine how many observations and variables are contained in the `countyComplete` data set. Write the answer in a complete sentence.

```{r}
str(county)
```

The `countyComplete` data frame contains 3116 observations (rows) and 56 variables (columns).

----


# Graphing Enhancements
This section will ask you to create graphics, often on data that needs pre-processing using the dplyr verbs you learned in the last homework. No one expects you to have all those commands memorized yet, the key is to know how to go find what you need. So reference last week's notes and homework while you do this section. 

## Exploring the murder rate 
This section demonstrates how you can build a nice graph in stages. Each question will have you adjust one thing on the plot. Instead of retyping the entire code each time, copy the code from the prior question and then make the requested adjustment. 

> This is the suggested way of building an advanced plot. Small bit by bit, changing or adding only one thing at a time. 

1. Using the `crime` data, create a density plot of the distribution of `murder` rate (x) by `region` (fill). To avoid problems with overplotting, use `alpha=.5` inside the density geometry to make the colors transparent. 

```{r}
ggplot(crime, aes(x=murder, fill=region)) + geom_density(alpha=.5)
```

2. These are pretty diverse regions, let's see how the murder rate varies within region. Fill by `division` instead, and panel on `region`. 
```{r}
ggplot(crime, aes(x=murder, fill=division)) + geom_density(alpha=.5) + facet_wrap(~region)
```

3. Add a different theme to this plot. 

```{r}
ggplot(crime, aes(x=murder, fill=division)) +
      geom_density(alpha=.5) + facet_wrap(~region) + 
      theme_bw()
```

4. Change the colors to use the `Set1` palette. 

```{r}
ggplot(crime, aes(x=murder, fill=division)) +
      geom_density(alpha=.5) + facet_wrap(~region) + 
      theme_bw() + scale_fill_brewer(palette="Set1")
```

5. Capitalize the legend title. _Hint, you're adding something to the `scale_` section_. 
```{r}
ggplot(crime, aes(x=murder, fill=division)) +
      geom_density(alpha=.5) + facet_wrap(~region) + 
      theme_bw() + scale_fill_brewer(name="Division", palette="Set1")
```

6. This works, but i'm not entirely thrilled with how much space it's taking up. Let's look at these as boxplots instead of density plots. _Don't forget that ggplot wants the continuous variable on the y axis_. 
```{r}
ggplot(crime, aes(y=murder, fill=division)) +
      geom_boxplot(alpha=.5) + facet_wrap(~region) + 
      theme_bw() + scale_fill_brewer(name="Division", palette="Set1")
```

7. The x-axis is all weird now. Let's be clear that the x is division. 
```{r}
ggplot(crime, aes(y=murder, x=division, fill=division)) +
      geom_boxplot(alpha=.5) + facet_wrap(~region) + 
      theme_bw() + scale_fill_brewer(name="Division", palette="Set1")
```

8. In this case, Division is nested within Region and so we don't need ALL the levels of division to show up in each panel. Let's "Free" our x axis by adding `scales="free_x"` to the `facet_wrap`. 

```{r}
ggplot(crime, aes(y=murder, x=division, fill=division)) +
      geom_boxplot(alpha=.5) + facet_wrap(~region, scales="free_x") + 
      theme_bw() + scale_fill_brewer(name="Division", palette="Set1")
```


9. But I really want to see them as horizontal plots, so add a `+ coord_flip()` layer. This also helps with readability of long axis names. 
```{r}
ggplot(crime, aes(y=murder, x=division, fill=division)) +
      geom_boxplot(alpha=.5) + facet_wrap(~region, scales="free") + 
      theme_bw() + scale_fill_brewer(name="Division", palette="Set1") + 
      coord_flip()
```

10. The legend is kinda useless now because we have the labels clearly written. Add `guide=FALSE` to the `scale` layer. 
```{r}
ggplot(crime, aes(y=murder, x=division, fill=division)) +
      geom_boxplot(alpha=.5) + facet_wrap(~region, scales="free") + 
      theme_bw() + scale_fill_brewer(name="Division", palette="Set1", guide=FALSE) + 
      coord_flip()
```

11. Add `ncol=1` to the `facet_wrap` to stack the panels vertically. 

```{r}
ggplot(crime, aes(y=murder, x=division, fill=division)) +
      geom_boxplot(alpha=.5) + facet_wrap(~region, scales="free", ncol=1) + 
      theme_bw() + scale_fill_brewer(name="Division", palette="Set1", guide=FALSE) + 
      coord_flip()
```


12. Add an appropriate title and axis titles. 
```{r}
ggplot(crime, aes(y=murder, x=division, fill=division)) +
      geom_boxplot(alpha=.5) + facet_wrap(~region, scales="free", ncol=1) + 
      theme_bw() + scale_fill_brewer(name="Division", palette="Set1", guide=FALSE) + 
      coord_flip() + ggtitle("Murder rate by region and division") + ylab("Division") +
      xlab("Murder Rate")
```



