---
title: "Homework 2"
author: "ENTER STUDENT ID"
date: "DATE"
output:
  pdf_document: default
  html_document: default
---

# Introduction

In this homework assignment, you will be working with data and managing factors. The data sets you will use are, `ncbirths` and `smoking` data set which come part of the `openintro` package.

The code chunk below sets some code chunk options (using `opts_chunk` from the `knitr` package) to make your knitted report output more readable. I encourage you to play around with these options to learn how they work. 

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.align='center') 
library(dplyr)
library(forcats)
ncbirths <- openintro::ncbirths
smoking <- openintro::smoking
```

# Working with Data

This section we will use `ncbirths` data set. 

1. Calculate the mean age of the mothers (`mage`).
```{r}

```

2. Pregnancies last on average 38 weeks. Recode the `weeks` variable to change all records where `weeks` is greater than 38, to equal 38. That is, for all record where `weeks>38`, change the value of `weeks` to `<- 38`.

```{r}

```

3. Use the `summary` function to calculate summary statistics on the fathers age (`fage`). Round to 3 digits using the `digits=` argument. Don't forget that you can look at the bottom of the help for summary (`?summary`) file for examples on how to use this function.  
```{r}

```

4. Use the `is.na()` function to create a new variable called `missing_gained` on the `ncbirths` data set that identifies if data on the `gained` variable is missing.

```{r}

```
   
5. What percent of records are missing data on weight gained during pregnancy? There are several ways you can calculate this! Explain what your process, and answer the question in a full sentence. 

```{r}

```


6. Use the `ifelse()` function to dichotomize the `weeks` variable at it's mean where records with values over the mean are labeled `AboveAve` and records with values below the mean are labeled `UnderAve`. Call this new variable `week_ave`. (_Hint: Calculate the mean value for the variable `weeks`, then use that number in the logical statement part of the `ifelse` function._)

```{r}

```

7. Create a frequency table for your new variable (`week_ave`) in the previous question. Then use the pipe operator `%>%` to add on the function `prop.table()` at the end. What does the `prop.table()` function do? 

```{r}

```


# Wrangling Factors 

In this section, you will use `smoking` data set.

1. Examine the variable `ethnicity` using `fct_count`. Then, collapse levels of that variable into a smaller number of factors using `fct_collapse`. _Hint: Create a new factor variable as part of the `smoking` data set (i.e smoking$ethnicity_new)._

* "NA" = c("Refused","Unknown")
* Asian = c("Chinese", "Asian")

```{r}

```

2. Create a two-way `table` of `ethnicity` against `ethnicity_new` to confirm that this new factor variable was created correctly.

```{r}

```

3. Using `fct_recode`, create a new factor variable `recode_ethnicity` from `ethnicity_new` with labels "A"(Asian), "B"(Black), "M"(Mixed), "W"(White). Make sure you create this new variable as part of the `smoking` data set. (i.e. `smoking$recode_ethnicity <-`) 

```{r}

```


4.  Manually reorder the level of `ethnicity_new` variable in an increasing order using `fct_relevel`.

```{r}

```


# Challenge Questions (1 point extra credit)

1. Within the `smoking` data set, define a new variable called `nationality_new` that collapses the levels "Unknown" and "Refused" from the `nationality` variable into a single level named "NA". Manually reorder the `nationality_new` variable in increasing order, then print a table with two variables: `nationality_new` and `smoke`.
_Hint: You can put two arguments in a table function. See ?table for details._

```{r}

```


