---
title: "Lesson 04 - Getting Started with R (notes)"
author: "PUT YOUR NAME HERE"
output: html_document
---

# Creating objects in R

```{r}
weight_kg <- 55
```

When assigning a value to an object, R does not print anything. You can force R to print the value by using parentheses or by typing the object name:

```{r}
weight_kg <- 55    
(weight_kg <- 55)  
weight_kg          
```

For instance, we may want to convert this weight into pounds (weight in pounds is 2.2 times the weight in kg):

```{r}
2.2 * weight_kg
```

We can also change an object's value by assigning it a new one:

```{r}
weight_kg <- 57.5
2.2 * weight_kg
```

For example, let's store the animal's weight in pounds in a new object, `weight_lb`:

```{r}
weight_lb <- 2.2 * weight_kg
```

and then change `weight_kg` to 100.

```{r}
weight_kg <- 100
```

R executes code in top-down order. So what happens on line 10 occurs before line 11. What do you think is the current content of the object `weight_lb`? 126.5 or 220?

> write your answer here


----

# Functions and their arguments

```{r}
sqrt(4)
```

Let's look into the `round` function. 

```{r}
round(3.14159)
```

We see that if we want a different number of digits, we cantype `digits = 2` or however many we want.

```{r}
round(3.14159, digits = 2)
```

If you provide the arguments in the exact same order as they are defined you
don't have to name them:

```{r}
round(3.14159, 2)
```

And if you do name the arguments, you can switch their order:

```{r}
round(digits = 2, x = 3.14159)
```

----

# Data Types

R objects come in different data types. 

## Numbers

```{r}
im_a_number <- 50
im_a_number*2
```

## Letters

```{r, error=TRUE}
(im_a_character <- "dog")
im_a_character*2
```

## Boolean

```{r}
3>4
sqrt(4)==2
```

# Data Structures

## Vectors 

For example we can create a vector of animal weights and assign it to a new object `weight_g`:

```{r}
(weight_g <- c(50, 60, 65, 82))
```

A vector can also contain characters:

```{r}
(animals <- c("mouse", "rat", "dog"))
```

You can use the function `class()` to see what data type a vector is. 

```{r}
class(weight_g)
class(animals)
```

If you combine letters and numbers, everything will be treated as a number. 
```{r}
(mix_match <- c(weight_g, animals))
class(mix_match)
```



## Doing math on vectors

You can perform math operations on the elements of a vector such as
```{r}
weight_KG <- weight_g/1000
weight_KG
```

When adding two vectors together, the elements in the same position are added to each other. So element 1 in the vector `a` is added to element 1 in vector `b`. 
```{r}
a <- c(1,2,3)
b <- c(6,7,8)
a+b
```

More complex calculations can be performed on multiple vectors. 
```{r}
wt_lb <- c(155, 135, 90)
ht_in <- c(72, 64, 50)
bmi <- 703*wt_lb / ht_in^2
bmi
```
      
All these operations on vectors behave the same way when dealing with variables in a data set (data.frame).   
   
If you want to add the values _within_ a vector, you use functions such as `sum()`, `max()` and `mean()`
```{r}
sum(a)
max(b)
mean(a+b)
```

## Subsetting vectors

If we want to extract one or several values from a vector, we must provide one
or several indices in square brackets. For instance:

```{r}
animals <- c("mouse", "rat", "dog", "cat")
animals[2]
animals[c(2, 3)]
```

The number in the indices indicates which element to extract. For example we can extract the 3rd element in `weight_g` by typing

```{r}
weight_g[3]
```

### Conditional subsetting

Typically, these logical vectors are not typed by hand, but are the output of other functions or logical tests such as: 

```{r}
weight_g > 50
```

For instance, if you wanted to select only the values where weight in grams is above 50 we would type:

```{r}
weight_g[weight_g > 50]
```

You can combine multiple tests using `&` (both conditions are true, AND) or `|` (at least one of the conditions is true, OR):

_Weight is less than 30g or greater than 60g_
```{r}
weight_g[weight_g < 30 | weight_g > 60]
```
_Weight is between 60 and 80lbs_
```{r}
weight_g[weight_g >= 60 & weight_g <= 80]
```

The function `%in%` allows you to test if any of the elements of a search vector are found:

```{r}
animals <- c("mouse", "rat", "dog", "cat")
animals[animals == "cat" | animals == "rat"] # returns both rat and cat
animals %in% c("rat", "cat", "dog", "duck", "goat")
animals[animals %in% c("rat", "cat", "dog", "duck", "goat")]
```

## Order matters. 
When considering string or character vectors or data elements, R treats everything in alphabetical order. 

```{r}
"four" > "five"
```

# Data Frames
We can load the `diamonds` data set into our global environment by typing 

```{r}
diamonds <- ggplot2::diamonds
```

We can get an idea of the structure of the data frame including variable names and types by using the `str` function,
```{r}
str(diamonds)
```

## Inspecting data.frame objects

## Identifying variables

Below is an example of finding the average price for all diamonds in the data set. 

```{r}
mean(diamonds$price)
```

Here is an example of finding the average price for `Good` quality diamonds. 
```{r}
mean(diamonds$price[diamonds$cut=="Good"])
```
