---
title: "Lesson 06 - Controlling your factors (notes)"
output: html_document
---

This notes file is a place for you to practice writing and executing the code found in the lecture notes. This is also a spot where you can (and should) write your own notes and thoughts. Explain what you are doing in each code chunk in your own words. 


## Preparation
Before class, run the following code chunk to make sure it works. 
If it does not, try to understand what the error message is telling you. Refer to the [FAQ page](https://norcalbiostat.github.io/MATH130/faq.html) for assistance. 

```{r}
library(forcats)
email <- openintro::email
```

----

# Introduction



# What is a factor?

First we should confirm that R sees `number` as a factor. 
```{r}

```

We can use the `levels()` function to get to know factor variables. 

```{r}

```

How many records are in each level?

----

# Convert a number to Factor

**Example: Is the email flagged as spam?**

Use the function `factor()` to convert the numeric variable `spam` to a factor, applying `lables` to convert 0 to "no" and 1 to "yes". 

```{r}

```

**Always confirm your recode** 

```{r}

```

----

# Factor ordering


## Manually specified 

```{r}

```


-----

# Factor (re)naming


```{r}

```


# Collapsing factor levels. 


## Chaining 


----

# Removing factor levels

