---
title: "Lesson 10 - Graphing Enhancements"
output: html_document
---

# Introduction 

```{r, warning=FALSE, message=FALSE}
library(ggplot2); library(dplyr)
knitr::opts_chunk$set(warning=FALSE, message=FALSE, error=TRUE) 
email <- read.table("../data/email.txt", header=TRUE, sep="\t")
NCbirths <- read.csv("../data/NCbirths.csv", header=TRUE)
```

# Appearance enhancements

## Themes

Try several themes. 

```{r}
ggplot(email, aes(x=line_breaks, y=num_char)) + geom_point() + theme_
ggplot(email, aes(x=line_breaks, y=num_char)) + geom_point() + theme_
ggplot(email, aes(x=line_breaks, y=num_char)) + geom_point() + theme_
```

## Legend modifications

```{r}
ggplot(email, aes(y=num_char, x=number, fill=number)) + geom_boxplot() + 
  scale_fill_discrete(name="________")
```

Color the points by a discrete variable
```{r}
ggplot(email, aes(y=num_char, x=number, col=number)) + geom_boxplot() + 
  scale________discrete(name="Size of number")
```


```{r}
ggplot(email, aes(y=num_char, x=number, col=number)) + geom_boxplot() + 
  scale_color_discrete(name="Size of number") + theme(_______)
```

# Removing NA categories. 

```{r}
NCbirths %>% _____(mage, whitemom, marital, habit) %>% ______() %>% 
  _____(aes(x=mage, fill=whitemom)) + 
  geom_density(alpha=.2) + facet_grid(marital~habit)
``

## Add means to boxplots. 

```{r}
ggplot(email, aes(y=num_char, x=number, fill=______)) + geom_boxplot() + 
  scale_fill_discrete(name="Size of number") + 
  stat_summary(fun.y="mean", geom="point", size=3, pch=17,color="___")
```


# Changing colors

## Manual

Using names or hex codes.

```{r}
ggplot(email, aes(x=number, fill=number)) + geom_bar() +
    scale_fill_manual(values=c("_______", "_______", "_______"))
```
 
 
 
## Using a color palette (recommended)
```{r}
library(RColorBrewer)
ggplot(email, aes(x=number, fill=number)) + geom_bar() +
    scale_fill_brewer(palette="_____")
```


# Barchart enhancements

## Numbers to the top of bars. 

```{r}
ggplot(email, aes(x=number)) + theme_bw() + 
    geom_bar(aes(y = ..count..)) + ggtitle("Frequency of emails by size of number") + 
    geom_text(aes(y=..count.. + 50, label=..count..), stat='count', size = 5)
```

## Plotting Proportions

```{r}
library(_______)

plot_frq(email$number)
```

Grouped proportions: 

```{r}
______(NCbirths$mature, NCbirths$marital)
```

Change to row proportions and remove the total bar. 
```{r}
_______(NCbirths$mature, NCbirths$marital, margin=_____, __________ = FALSE)
```

# Multivariate Plotting 

## Boxplots with three variables. 

```{r}
ggplot(NCbirths, aes(y=______, x=_____, fill=______)) + 
        geom_boxplot()
```

## Faceting

```{r}
ggplot(NCbirths, aes(x=mage, fill=gender)) + geom_density(alpha=.3) + facet_wrap(~____)
```

Using the `ncol` argument in the `facet_wrap()` statement. 
```{r, fig.height=6}
ggplot(NCbirths, aes(x=mage, fill=gender)) + geom_density(alpha=.3) + ________(~____, ncol=1)
```
 

# Multiple plots per window

#### ggplot

```{r}
library(_______)
_____ <- ggplot(NCbirths, aes(x=mage, fill=whitemom)) + geom_density(alpha=.2)
_____ <- ggplot(NCbirths, aes(x=mage, col=whitemom)) + geom_density() 
grid.arrange(_____,_____, ncol=2)
```


## Paneling on two variables

```{r, fig.width=10, fig.height=5}
ggplot(NCbirths, aes(x=mage, fill=whitemom)) + geom_density(alpha=.2) + _______(marital~habit)
```



